---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# protoclust

This R package implements hierarchical clustering with minimax linkage (a.k.a., prototype clustering).  See
[Bien, J. and Tibshirani, R. (2011) "Hierarchical Clustering with Prototypes via Minimax Linkage"](http://faculty.bscb.cornell.edu/~bien/papers/jasa2011minimax.pdf) for more details.

Since `protoclust` is on CRAN, it can be easily installed using
```{r, eval = FALSE}
install.packages("protoclust")
```
in R.  Occasionally, this github version of the package will be more up-to-date.  The easiest way to install this version is
by using the [devtools](https://cran.r-project.org/web/packages/devtools/index.html) R package (if not already installed, open R and type `install.packages("devtools")`). To install `protoclust`, type

```{r, eval = FALSE}
devtools::install_github("jacobbien/protoclust")
```
in R.

## Simple example of prototype clustering

Suppose we have some data.
```{r data, echo = FALSE}
set.seed(1)
n <- 100
p <- 2
x <- matrix(rnorm(n * p), n, p)
rownames(x) <- paste("A", 1:n, sep = "")
d <- dist(x)
par(pty = "s")
plot(x, type = "n", xlab = "", ylab = "")
points(x, pch = 20, col = "lightblue")
```

We compute an n-by-n matrix of dissimilarities `d`.

```{r dendrogram}
# perform minimax linkage clustering:
library(protoclust)
hc <- protoclust(d)

# cut the tree at height 1:
h <- 1
cut <- protocut(hc, h = h)
# plot dendrogram (and show cut):
plot(hc, imerge = cut$imerge, col=2)
abline(h = h, lty = 2)
```

This gives us 9 prototypes with the guarantee that each of the n points is no more than 1 unit from one of the prototypes.

```{r dataclusters, echo = FALSE}
# get the prototype assigned to each point:
pr <- cut$protos[cut$cl]
# find point farthest from its prototype:
dmat <- as.matrix(d)
ifar <- which.max(dmat[cbind(1:n, pr[1:n])])

# since this is a 2d example, make 2d display:
par(pty = "s")
plot(x, type="n", xlab = "", ylab = "")
points(x, pch=20, col="lightblue")
lines(rbind(x[ifar, ], x[pr[ifar], ]), col=3)
points(x[cut$protos, ], pch=20, col="red")
text(x[cut$protos, ], labels=hc$labels[cut$protos], pch=19)
tt <- seq(0, 2 * pi, length=100)
for (i in cut$protos) {
 lines(x[i, 1] + h * cos(tt), x[i, 2] + h * sin(tt))
}
```

